Detailed Geant4 simulations of the ANITA and ANITA-CUP neutron facilities

### 2015-04-17 Paper2015.tex
Edited bibliography and citations.

\section{Introduction}
\todo{Review introduction}
\IEEEPARstart{M}{onte} Carlo simulations of naked spallation neutron sources at Los Alamos Neutron Science Center (LANSCE) Weapons Neutron Research (WNR) Target 4\cite{Wender87} and the ANITA beam at the University of Uppsala The Svedberg Laboratory (TSL) have been completed using Geant4 and proved that Binary intranuclear cascade (INC) model gave a good representation of neutron spectra than Bertini INC model did\cite{Platt13}. Simulation result of neutron spectra using Geant4 with Binary INC model is below measurement data from ANITA neutron facility. It is likely to be influenced by our omission of collimator, shielding, and bending magnet component. This work is to verify this hypothesis is correct. In addition, we are intested in the gamma field at CUP and SUP in spallation neutron source at the ANITA neutron facility at TSL for SEE testing.

\section{Modelling}
Simulations were conducted using Geant4 version 10.0 and the binary intranuclear cascade model, as recommended in~\cite{Platt13}.
A detailed model of the ANITA geometry was implemented in Geant4, as illustrated in Fig.~\ref{fig:ANITAoverview}
Description of the geometry is given in \cite{Prokofiev2009,Prokofiev14}.
%; a summary only is given here.
%The spallation target is a tungsten cylinder of diameter \SI{5}{\cm} and length \SI{2.4}{\cm}~\cite{ANITAdrawing}.
%The target is cooled by water and surrounded by a stainless-steel cooling jacket.
%The target assembly with its cooling jacket appears as a small blue rectangle in Fig.~\ref{fig:ANITAoverview}.
%Lead shielding blocks in the target region are also modelled, and shown in red Fig.~\ref{fig:ANITAoverview}.
%A large electromagnet surrounding the target is modelled and shown in Fig.~\ref{fig:ANITAoverview} in grey (iron) and yellow (copper).
%An iron collimator with variable apertures is modelled  and shown downstream of the target (grey).
Except where indicated otherwise, all results in this summary are for the standard collimator size (\SI{10.2}{\cm} diameter).
In the final paper we will present results for cylindrical collimators in the range from \SIrange{3}{30}{\cm} diameter.
Simulated protons were incident axially at an energy of \SI{180}{\MeV}.
Resulting gamma and neutron fields were evaluated at several detector locations.
In this paper results are presented for three of these: the Standard User Position (SUP), \SI{2.5}{\m} downstream of the target~\cite{Prokofiev2009}, the Close User Position (CUP), \SI{0.75}{\m} downstream of the target, and the CUP-TOF, \SI{0.84}{m} downstream of the target.
%N.B. This is downstream of the downstream edge of the target in Qian's model.
The SUP and CUP positions are of interest as they are positions used for SEE tests; the CUP-TOF position is the location of thin-film breakdown counter (TFBC) detectors used for beam monitoring and characterisation using time-of-flight (TOF) techniques.
Model validation will be demonstrated against CUP-TOF data (section~\ref{tbd}).

@@

\section{Results}
In previous work Platt et al.~\cite{Platt13} reported simulations of a naked ANITA target, demonstrating good qualitative agreement with independent calculations and measurements of the fast neutron field at ANITA SUP~\cite{Prokofiev2009}.
Part of the motivation for this work was to investigate whether improvements in the fidelity of the geometry used in the Geant4 model would lead to increased quantitative agreement.
\todo{Review first part of Results section -- move to introduction?}

\subsection{Neutrons}
Fig.~\ref{fig:SUPDensity} shows neutron spatial distribution at the SUP, for neutrons above \SI{10}{\MeV}.
The collimation effect is clearly visible.
Within the beam umbra the predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{7.1e5}{\neutron\per\cm\squared\per\second}, compared to \SI{5.7e5}{\neutron\per\cm\squared\per\second} from earlier calculations~\cite{Platt13} and \SI{9.3e5}{\neutron\per\cm\squared\per\second} from an analytical model~\cite{Prokofiev2009}.
The analytical model therefore exceeds the present calculations by 31\%.

@@

Fig.~\ref{fig:CUPDensity} shows equivalent data for the CUP.
In the central region the predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{0}{\neutron\per\cm\squared\per\second}, compared to \SI{0}{\neutron\per\cm\squared\per\second} from an analytical model~\cite{Prokofiev14}.
The analytical model therefore exceeds the present calculations about TBD\%.
\todo{Complete quantification at CUP}.
The results show\ldots
\todo{Describe the results: contours and profile at CUP and SUP}

Fig.~\ref{fig:CUPProfile} shows neutron fluence profile at the CUP, compared with measurements made with \U238\ TFBCs~\cite{Prokofiev14}.
In Fig.~\ref{fig:CUPProfile} simulated data were folded in energy with the neutron fission cross section for \U238~\cite{tbd}.
The results show\ldots
\todo{Describe the results: contours and profile at CUP}

@@

Fig.~\ref{fig:Lethargyplots} compares calculated spectra at SUP and CUP with their respective analytical models~\cite{Prokofiev2009,Prokofiev14}.
The SUP data are also compared with results from preliminary Geant4 simulations~\cite{Platt13}.
The results show\ldots
\todo{Describe the neutron spectra.}

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{SUPNormalisedRADECS.eps}
        \subcaption{
            SUP.
            The comparison is against preliminary Geant4 modelling with a naked target model~\cite{Platt13} and against an analytical model~\cite{Prokofiev2009}
        }
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{ANITALetYieldnormalisedRADECS.eps}
        \subcaption{
            CUP.
            The comparison is against the a standard analytical model~\cite{Prokofiev2009}
        }
    \end{minipage}
	\caption{
        Calculated neutron spectra, normalised to fluence above \SI{10}{\MeV}.
        \SI{10.2}{\cm} diameter collimator
    }
	\label{fig:Lethargyplots}
\end{figure}

Fig.~\ref{fig:DifferentialSpectra} summarises neutron spectra results by comparing differential flux as calculated in this work with analytical models~\cite{Prokofiev2009,Prokofiev14}.
The SUP calculations and analytical models agree very closely below \SI{20}{\MeV}; the analytical model exceeds the calculations somewhat above that energy.
At CUP the analytical model exceeds the Geant4 calculations by about TBD\% below \SI{20}{\MeV} and slightly more above that energy.
\todo{Improve precision of the summary description of differential spectra.}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\columnwidth]{DiffYieldComparedSUPCUP10.eps}
    \caption{
        Differential neutron spectra at CUP and SUP (\SI{10.2}{\cm} diameter collimator), comparing results of this work to analytical models~\cite{Prokofiev2009,Prokofiev14}
    }
    \label{fig:DifferentialSpectra}
\end{figure}

@@

\bibitem{Wender87}
S.A. Wender and P.W. Lisowski,
\newblock A white neutron source from 1 to 400 MeV,										% title	
\newblock {\em Nucl. Instr. Meth. Phys. Res. B}, vol. 24-25, pp. 897-900.					% vol:page, 2011

\bibitem{Platt13}
Platt, S.P. and Hong, Q. and Mein, S.J. and Zhang, L.H.,
\newblock Neutron and gamma fields at neutron spallation sources for single-event-effects testing,		% title	
\newblock {\em in Proc. 14th European Conf. Radiat. Effects Compon. Syst. 2013,} pp 1-4			% vol:page, 2011

\bibitem{Prokofiev14}
Prokofiev, A.V. and Blomgren, J. and Majerle, M. and Nolte, R. and Rottger, S. and Platt, S.P. and Cai Xiao Xiao and Smirnov, A.N.,
\newblock Characterization of the ANITA Neutron Source for Accelerated SEE Testing at the Svedberg Laboratory,		% title	
\newblock {\em in Proc. IEEE Radiation Effects Data Workshop. 2009,} pp 166-173 									% vol:page, 2011

\bibitem{Prokofiev2009}
Prokofiev, A.V. and Passoth, E. and Hjalmarsson, A. and Majerle, M.,
\newblock CUP-A New High-Flux Irradiation Position at the ANITA Neutron Facility at TSL,		% title	
\newblock {\em in Proc. IEEE Trans. Nucl. Sci. 2014}, pp 1929-1936 							% vol:page, 2011

\bibitem{ANITAdrawing}
Prokofiev, A.V.
\newblock personal communication
\newblock {\em TSL, Sep. 2013}

>>>>

\section{Introduction}
\todo{Review introduction}
\IEEEPARstart{M}{onte} Carlo simulations of naked spallation neutron sources at Los Alamos Neutron Science Center (LANSCE) Weapons Neutron Research (WNR) Target 4\cite{Wender87} and the ANITA beam at the University of Uppsala The Svedberg Laboratory (TSL) have been completed using Geant4 and proved that Binary intranuclear cascade (INC) model gave a good representation of neutron spectra than Bertini INC model did\cite{Platt2013}. Simulation result of neutron spectra using Geant4 with Binary INC model is below measurement data from ANITA neutron facility. It is likely to be influenced by our omission of collimator, shielding, and bending magnet component. This work is to verify this hypothesis is correct. In addition, we are intested in the gamma field at CUP and SUP in spallation neutron source at the ANITA neutron facility at TSL for SEE testing.

\section{Modelling}
Simulations were conducted using Geant4 version 10.0 and the binary intranuclear cascade model, as recommended in~\cite{Platt2013}.
A detailed model of the ANITA geometry was implemented in Geant4, as illustrated in Fig.~\ref{fig:ANITAoverview}
Description of the geometry is given in \cite{Prokofiev2009,Prokofiev2014}.
%; a summary only is given here.
%The spallation target is a tungsten cylinder of diameter \SI{5}{\cm} and length \SI{2.4}{\cm}~\cite{ANITAdrawing}.
%The target is cooled by water and surrounded by a stainless-steel cooling jacket.
%The target assembly with its cooling jacket appears as a small blue rectangle in Fig.~\ref{fig:ANITAoverview}.
%Lead shielding blocks in the target region are also modelled, and shown in red Fig.~\ref{fig:ANITAoverview}.
%A large electromagnet surrounding the target is modelled and shown in Fig.~\ref{fig:ANITAoverview} in grey (iron) and yellow (copper).
%An iron collimator with variable apertures is modelled  and shown downstream of the target (grey).
Except where indicated otherwise, all results in this summary are for the standard collimator size (\SI{10.2}{\cm} diameter).
In the final paper we will present results for cylindrical collimators in the range from \SIrange{3}{30}{\cm} diameter.
Simulated protons were incident axially at an energy of \SI{180}{\MeV}.
Resulting gamma and neutron fields were evaluated at several detector locations.
In this paper results are presented for three of these: the Standard User Position (SUP), \SI{2.5}{\m} downstream of the target~\cite{Prokofiev2009}, the Close User Position (CUP), \SI{0.75}{\m} downstream of the target, and the CUP-TOF, \SI{0.84}{m} downstream of the target.
%N.B. This is downstream of the downstream edge of the target in Qian's model.
The SUP and CUP positions are of interest as they are positions used for SEE tests; the CUP-TOF position is the location of thin-film breakdown counter (TFBC) detectors used for beam monitoring and characterisation using time-of-flight (TOF) techniques.
Model validation will be demonstrated against CUP-TOF data (section~\ref{tbd}).

@@

\section{Results}
In previous work Platt et al.~\cite{Platt2013} reported simulations of a naked ANITA target, demonstrating good qualitative agreement with independent calculations and measurements of the fast neutron field at ANITA SUP~\cite{Prokofiev2009}.
Part of the motivation for this work was to investigate whether improvements in the fidelity of the geometry used in the Geant4 model would lead to increased quantitative agreement.
\todo{Review first part of Results section -- move to introduction?}

\subsection{Neutrons}
Fig.~\ref{fig:SUPDensity} shows neutron spatial distribution at the SUP, for neutrons above \SI{10}{\MeV}.
The collimation effect is clearly visible.
Within the beam umbra the predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{7.1e5}{\neutron\per\cm\squared\per\second}, compared to \SI{5.7e5}{\neutron\per\cm\squared\per\second} from earlier calculations~\cite{Platt2013} and \SI{9.3e5}{\neutron\per\cm\squared\per\second} from an analytical model~\cite{Prokofiev2009}.
The analytical model therefore exceeds the present calculations by 31\%.

@@

Fig.~\ref{fig:CUPDensity} shows equivalent data for the CUP.
In the central region the predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{0}{\neutron\per\cm\squared\per\second}, compared to \SI{0}{\neutron\per\cm\squared\per\second} from an analytical model~\cite{Prokofiev2014}.
The analytical model therefore exceeds the present calculations about TBD\%.
\todo{Complete quantification at CUP}.
The results show\ldots
\todo{Describe the results: contours and profile at CUP and SUP}

Fig.~\ref{fig:CUPProfile} shows neutron fluence profile at the CUP, compared with measurements made with \U238\ TFBCs~\cite{Prokofiev2014}.
In Fig.~\ref{fig:CUPProfile} simulated data were folded in energy with the neutron fission cross section for \U238~\cite{tbd}.
The results show\ldots
\todo{Describe the results: contours and profile at CUP}

@@

Fig.~\ref{fig:Lethargyplots} compares calculated spectra at SUP and CUP with their respective analytical models~\cite{Prokofiev2009,Prokofiev2014}.
The SUP data are also compared with results from preliminary Geant4 simulations~\cite{Platt2013}.
The results show\ldots
\todo{Describe the neutron spectra.}

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{SUPNormalisedRADECS.eps}
        \subcaption{
            SUP.
            The comparison is against preliminary Geant4 modelling with a naked target model~\cite{Platt2013} and against an analytical model~\cite{Prokofiev2009}
        }
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{ANITALetYieldnormalisedRADECS.eps}
        \subcaption{
            CUP.
            The comparison is against the a standard analytical model~\cite{Prokofiev2009}
        }
    \end{minipage}
	\caption{
        Calculated neutron spectra, normalised to fluence above \SI{10}{\MeV}.
        \SI{10.2}{\cm} diameter collimator
    }
	\label{fig:Lethargyplots}
\end{figure}

Fig.~\ref{fig:DifferentialSpectra} summarises neutron spectra results by comparing differential flux as calculated in this work with analytical models~\cite{Prokofiev2009,Prokofiev2014}.
The SUP calculations and analytical models agree very closely below \SI{20}{\MeV}; the analytical model exceeds the calculations somewhat above that energy.
At CUP the analytical model exceeds the Geant4 calculations by about TBD\% below \SI{20}{\MeV} and slightly more above that energy.
\todo{Improve precision of the summary description of differential spectra.}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\columnwidth]{DiffYieldComparedSUPCUP10.eps}
    \caption{
        Differential neutron spectra at CUP and SUP (\SI{10.2}{\cm} diameter collimator), comparing results of this work to analytical models~\cite{Prokofiev2009,Prokofiev2014}
    }
    \label{fig:DifferentialSpectra}
\end{figure}

@@

\bibitem{Prokofiev2009}
Prokofiev, A. V. et al., 
\newblock CUP-A New High-Flux Irradiation Position at the ANITA Neutron Facility at TSL,		\newblock {\em Proc. IEEE Trans. Nucl. Sci. 2014}, pp 1929-1936

\bibitem{Prokofiev2014}
Prokofiev, A. V. et al.,
\newblock Characterization of the ANITA Neutron Source for Accelerated SEE Testing at the Svedberg Laboratory,
\newblock {\em in Proc. IEEE Radiat. Effects Data Work. 2009,} pp 166-173

\bibitem{Platt2013}
Platt, S. P. et al.,
\newblock Neutron and gamma fields at neutron spallation sources for single-event-effects testing,
\newblock {\em in Proc. 14th Euro. Conf. Radiat. Effects Compon. Syst. 2013}

\bibitem{tbd0} \todo{Add referencefor Geant4}

\bibitem{tbd1} \todo{Complete reference for \U238}

\bibitem{tbd2} \todo{Complete reference for dose conversion}
